Understanding the Nanoscale Structure of Inverted Hexagonal Phase Lyotropic Liquid Crystal Polymer Membranes

### 2018-11-18 Ben_Manuscripts/structure_paper/response_to_reviewers.tex

	%BJC: want to convery that we want to reproduce all of the experimental features as best as possible
	% using an out-of-the-box force field. That way we know we are approximating the chemical environment
	% as close to the experimental system as possible. The goal is to eventually test large numbers of 
	% monomers without the needing to dive into the excruciating detail shown here.

    \end{quote}

    \item \begin{quote}
    \textit{1. For the simulations performed here, the authors do make a strong case that 5 columns per
    core gets the best pore-to-pore spacing (Figure 5). However, insufficient time is accessed to see 
    whether the 4-, 5-, and 6-column initial conditions move towards or away from the expected pore 
    spacing over many-microseconds trajectories. That said, for the quenches performed here, 5-columns
    per core seems more plausible than 4.}
    \end{quote}

    Author reply: We agree with the reviewer that 5 columns-per-core is more plausible than 4 and that much
    longer simulations may provide further supporting evidence. However, there is very little motion on
    the hundreds of nanoseconds timescale that we simulated which suggests that we will not gain any 
    significant new insight by extending our simulations into the microsecond regime. It is not feasible 
    for us to carry out these simulations as it would consume our computational resources with little reward.
    %BJC: The following might apply:
    Some confusion may have been generated by agreement of the disordered basin 6 column-per-pore system
    with experimental pore spacing. As mentioned in the text, the vertical stacking distance between monomers
    in that system is far too large. The columns are `stretched' in the $z$-direction allowing the pores take
    up less space on the $xy$ plane and thus pack closer together.

@@

    Author reply: We thank the reviewer for pushing us to provide more significant physical insight based
    on the number of monomers per column. The number of columns-per-pore influences the pore radius which
    is important for determining the limits of size-exclusion-based separations. It also tells use the 
    density of interacting sites within the pore which may play a role in small molecule transport. We have
    added the following text to the main document at the end of Section~\ref{M-section:mon_per_pore} on 
    page _:

    ``The number of columns per pore dictates the density of interaction sites within each pore and the 
    pore radius. A higher number of interaction sites will surely play a role in transport of molecules
    that have an affinity for the monomer head groups. 
    %We measured the pore radius as the average distance from the pore center to each head group COM. 
    The pore radius increases with the number of columns per pore (See Figure~\ref{S-fig:pore_radius}. 
    This radius is likely related to the size of the molecules which can be excluded by this type of 
    membrane. We may be able to control LLC membrane pore size by selecting monomers that are stable
    in configurations with more columns-per-pore."

    \item \begin{quote}
    \textit{2. The authors make a pretty compelling case that water causes the R-double scattering feature
    by showing its absence when water is added, but that the R-double appears and persists for 200ns after
    the water makes its way into the crystal. However, the authors also show that nonuniformly spaced
    monomers can give rise to the R-Double feature, a seeming counterexample to the necessity of water.
    Is the R-double feature with water because the monomers are now irregularly spaced? What happens when
    the waters are removed and the scattering re-performed in Fig 17c(bottom)?}
    \end{quote}

    Author reply: We thank the reviewer for exposing a potentially unclear
    explanation of the origin of R-double. R-double appears in a dry system with
    non-uniformly spaced monomers \textit{only} if $z$-direction position
    restraints are applied to monomer head groups in order to maintain unevenly
    spaced stacking. As soon as the restraints are released, R-double fades and the
    monomer spacing becomes uniform. We modified the text on page _ in order to more
    clearly convey this message:

    Old text: ``We can also produce R-double if the LLC monomers are not
    uniformly spaced in the $z$-direction, but instead form pairs that stack less
    than 3.7~\AA~apart, with COMs that are spaced 7.4 \AA~ from neighboring pairs
    of monomers (Figure~\ref{fig:staggered_rzplot_norestraints}).  Our force field
    causes our system to tend towards uniformly spaced layers. Simulations of
    unevenly spaced systems are only stable if position restraints are applied to
    heavy atoms of the phenyl rings. Additionally, there is little evidence from QM
    studies of stacked $\pi-\pi$ systems that such uneven stacking could be
    energetically stable.~\cite{tauer_beyond_2005}"

    New text: ``We can also produce R-double if the LLC monomers are not
    uniformly spaced in the $z$-direction, but instead are placed in pairs that
    stack less than 3.7~\AA~apart, with COMs that are spaced 7.4 \AA~ from
    neighboring pairs of monomers (Figure~\ref{fig:staggered_rzplot_norestraints}).
    Simulations of unevenly spaced systems are only stable while position
    restraints are applied to heavy atoms of the phenyl rings. There is little
    evidence from quantum mechanical studies of stacked $\pi-\pi$ systems that such
    uneven stacking could be energetically stable.~\cite{tauer_beyond_2005} As soon
    as we remove position restraints, our system to tends towards uniformly spaced
    monomers."

    R-double is only long-term stable without position restraints if a small
    amount of water is present. We believe that water facilitates the appearance
    of R-double by stabilizing non-uniform spacing through hydrogen bonding, as the
    reviewer suggests. We clarified the text on page _ in order to make this 
    conclusion clear:

    Old text: ``When two vertically stacked monomer head groups hydrogen bond
    with a shared water molecule, the monomers are drawn closer together (as
    illustrated in Figure~\ref{fig:hbond_visualization}), which creates an
    asymmetry that allows R-double to appear. If a monomer head group shares a
    hydrogen-bonded water molecule with a head group above itself, it will be less
    likely to share a water molecule with a head group below it due to geometric
    constraints. The monomer head group below can just as easily share a water
    molecule with a head group below itself. In this scenario, the COMs of each
    pair are 2 times the $\pi$-stacking distance apart which would lead to R-double
    (much like the configuration in
     Figure~\ref{fig:staggered_rzplot_norestraints}). There are a modest number of
    occurrences of this scenario, which we quantify in further detail in the SI,
    Section \ref{S-section:rdouble}."

    New text: ``The addition of a small amount of water to the pores stabilizes
    non-uniform stacking of head groups much like that shown in
    Figure~\ref{fig:staggered_rzplot_norestraints}. When two vertically stacked
    monomer head groups hydrogen bond with a shared water molecule, the monomers
    are drawn closer together (as illustrated in
    Figure~\ref{fig:hbond_visualization}), which creates the asymmetry that allows
    R-double to appear. If a monomer head group shares a hydrogen-bonded water
    molecule with a head group above itself, it will be less likely to share a
    water molecule with a head group below it due to geometric constraints. The
    monomer head group below can just as easily share a water molecule with a head
    group below itself. There are a modest number of occurrences of this scenario,
    which we quantify in further detail in the SI, Section
    \ref{S-section:rdouble}."

    The reviewer asked us to re-perform the scattering analysis on the same
    system with water molecules removed. We would like to emphasize that we
    attempted to make a similar argument by removing just the head group atoms from
    the trajectory and re-performing the scattering analysis
    (Figure~\ref{M-fig:rdouble_nophenyls}). In the absence of head groups, R-double
    is not present, implying that their structure (which is influenced by the
    presence of water molecules) is responsible for the R-double reflection. For
    completeness, we performed the scattering analysis on the same system with
    water molecules removed, as suggested by the reviewer, and R-double is still
    present. This is the expected result since the structure of water itself does
    not give rise to R-double. We've added the resulting simulated XRD pattern to
    the Supporting Information, Figure~\ref{S-TBD}.

@@

    \item \begin{quote}
    \textit{3. The authors make the case that the R-spots feature comes from the structure of the hexagonal
    tails, not their tilt. The authors should perform the scattering analysis to check if the R-spots are
    present on the tilted-tails they constrain: This quick experiment would nail down their interpretation
    here, or show that the tilted tails are also plausible (but not observed in this model).}
    \end{quote}

    Author reply: We thank the reviewer for their suggestion. We do not deny
    that the tilt of the monomer tails is a plausible explanation for the
    appearance of reflections that resemble R-spots in many liquid crystal systems.
    While the reviewer's suggested experiment is a good idea, its results in this
    case are ambiguous for a couple reasons. First, the tails may still pack
    hexagonally in the tilted configuration which would independently give rise to
    R-spots. Second, it is necessary to impose high force constant (>1000 kJ mol
    nm$^{-2}$) position restraints in order to maintain a tilt angle close to that
    implied by the 2D WAXS data. The resulting high degree of order in the system 
    gives way to reflections which are easily misinterpretted. We have added the
    following text highlighted in red to the text shown on Page _ in order to
    acknowledge that tail tilt is a valid explanation of R-spot-like reflections 
    and to offer a reason a molecular reason why our monomers in our system do not
    tilt:

    Modified text: ``Previous literature has attributed the R-spots reflection
    in this particular WAXS dataset as the result of tilted alkane
    chains~\cite{feng_scalable_2014}. {\color{red} This explanation is not
    unfounded as there are examples in literature of systems where tilted liquid
    crystals give rise to reflections that resemble R-spots (Cite, Cite).} We
    measured the tilt angle of the alkane chains of the 280 K system by measuring
    the angle made by the vector extending from top to bottom of each tail with
    respect to the membrane plane.  We found that it equilibrates to an average
    tilt angle of -2 $\pm$ 13\degree~(Figure~\ref{fig:tilt}), far from the
    37\degree~tilt angle previously used to explain R-spots. Even when we placed
    position restraints on the monomers in a tilted initial configuration, monomers
    quickly reduced their tilt relative to the $xy$ plane to an angle statisically
    indistinguishable from zero (see Section \ref{S-section:tail_tilt} of the SI).
    {\color{red} While it is plausible that our force field may in part be
    responsible for the tendency of monomers to stay nearly flat, the monomer tails
    lack the functional groups which are known to facilitate tail tilting (cite).}"

@@

    \item \begin{quote}
    \textit{These arguments would be stronger if there weren't so many
	    qualifications about simulation timescales and GAFF, in particular its ability
	    to predict the shortest-scale feature (pi-stacking) for this system. How does
	    everything else change if the force-field were tuned to get that feature
	    correct?}
    \end{quote}

    Author reply: We thank the reviewer for their constructive criticism. Many
    of our findings may be relatively independent of the force fields used, but we
    feel that we should include these qualifications in order to not overstate our
    case. It is common for researchers to run simulations for an insufficient
    length of time and to make bold claims. We acknowledge the shortcomings of our
    techniques in order to inspire other researchers to build on our initial work.

    We agree with the reviewer that we could tune our force field in order to
    exactly reproduce the $\pi-\pi$ stacking distance seen experimentally. Surely
    there will be some structural differences. There are many ways to tune a force
    field in order to achieve this. We prefer to work with an "out-of-the-box"
    force field since they are widely used and can be applied to vast number of
    systems. The ultimate goal of our work is to create new monomers that can
    achieve specfic separations.  If we can gain experimentally-consistent insights
    using our model simulated with GAFF, then we can have increased confidence in
    future predictions generated from systems built with unsynthesized monomers.
    With that said, we think that it would be interesting to repeat our analysis
    using different and/or optimized force fields in a future study in order to
    comment on the signficance of any differences.

    %BJC: Emphasize main goal in intro (or pull a quote that says it)
    %BJC: add justification of gaff to methods?
    %BJC: Add force field optimization to future work

@@

    \item \begin{quote}
    \textit{The MD-Structure-Factor repository is not available at the provided
	    github link, nor does it appear to be online anywhere. So, the correctness of
	    that code cannot be checked.}
    \end{quote}

    Author reply: We thank the reviewer for pointing this out and apologize for
    our mistake. The referenced GitHub repository was set to private due to
    miscommunication with our collaborators but it has now been made public. Please
    feel free to read through the code. We have performed checks to ensure that the
    code works properly, especially when simulating monoclinic unit cells. For
    example, we generated test trajectories containing thousands of frames made up
    of configurations of randomly placed point scatterers in a monoclinic unit cell.
    We expect that the simulated diffraction pattern will have a gaussian shape.
    This exact result is shown in the Figure S-thingy of the supporting information.
    With improper treatment of the monoclinic geometry, the gaussian shape is
    skewed into an elliptical shape.

### 2018-11-18 Ben_Manuscripts/structure_paper/Final_Draft.tex
Added many responses to Reviewer 1. Outlined else needs to be done

  Previous literature has attributed the R-spots reflection in this particular
  WAXS dataset as the result of tilted alkane chains~\cite{feng_scalable_2014}.
  We measured the tilt angle of the alkane chains of the 280 K system by measuring
  the angle made by the vector extending from top to bottom of each tail with
  respect to the membrane plane. We found that it equilibrates to an average tilt
  angle of -2 $\pm$ 13\degree~(Figure~\ref{fig:tilt}), far from the 37\degree~tilt angle
  previously used to explain R-spots. Even when we placed position restraints on
  the monomers in a tilted initial configuration, monomers quickly reduced their
  tilt relative to the $xy$ plane to an angle statisically indistinguishable from
  zero (see Section \ref{S-section:tail_tilt} of the SI).

@@

  We can also produce R-double if the LLC monomers are not uniformly spaced in the
  $z$-direction, but instead form pairs that stack less than 3.7~\AA~apart, with
  COMs that are spaced 7.4 \AA~ from neighboring pairs of monomers 
  (Figure~\ref{fig:staggered_rzplot_norestraints}). 
  Our force field causes our system to tend towards uniformly spaced layers.
  Simulations of unevenly spaced systems are only stable if position restraints
  are applied to heavy atoms of the phenyl rings.  Additionally, there is little
  evidence from QM studies of stacked $\pi-\pi$ systems that such uneven stacking
  could be energetically stable.~\cite{tauer_beyond_2005}

@@

  When two vertically stacked monomer head groups hydrogen bond with a shared
  water molecule, the monomers are drawn closer together (as illustrated in
  Figure~\ref{fig:hbond_visualization}), which creates an asymmetry that allows
  R-double to appear. If a monomer head group shares a hydrogen-bonded water
  molecule with a head group above itself, it will be less likely to share a
  water molecule with a head group below it due to geometric constraints. The
  monomer head group below can just as easily share a water molecule with a head
  group below itself. In this scenario, the COMs of each pair are 2 times the
  $\pi$-stacking distance apart which would lead to R-double (much like the
  configuration in Figure~\ref{fig:staggered_rzplot_norestraints}). There are a
  modest number of occurrences of this scenario, which we quantify in further
  detail in the SI, Section \ref{S-section:rdouble}.

>>>>

  Previous literature has attributed the R-spots reflection in this particular
  WAXS dataset as the result of tilted alkane chains~\cite{feng_scalable_2014}.
  This explanation is not unfounded as there are examples in literature of
  systems where tilted liquid crystals give rise to reflections that resemble
  R-spots (Cite, Cite). We measured the tilt angle of the alkane chains of the
  280 K system by measuring the angle made by the vector extending from top to
  bottom of each tail with respect to the membrane plane. We found that it
  equilibrates to an average tilt angle of -2 $\pm$
  13\degree~(Figure~\ref{fig:tilt}), far from the 37\degree~tilt angle previously
  used to explain R-spots. Even when we placed position restraints on the
  monomers in a tilted initial configuration, monomers quickly reduced their tilt
  relative to the $xy$ plane to an angle statisically indistinguishable from zero
  (see Section \ref{S-section:tail_tilt} of the SI). While it is plausible that 
  our forcefield may in part be responsible for the tendency of monomers to stay
  nearly flat, the monomer tails lack the functional groups which are known to
  facilitate tail tilting.

@@

  We can also produce R-double if the LLC monomers are not uniformly spaced in
  the $z$-direction, but instead are placed in pairs that stack less than
  3.7~\AA~apart, with COMs that are spaced 7.4 \AA~ from neighboring pairs of
  monomers (Figure~\ref{fig:staggered_rzplot_norestraints}). Simulations of
  unevenly spaced systems are only stable while position restraints are applied
  to heavy atoms of the phenyl rings. There is little evidence from quantum
  mechanical studies of stacked $\pi-\pi$ systems that such uneven stacking could
  be energetically stable.~\cite{tauer_beyond_2005} As soon as we remove position
  restraints, our system to tends towards uniformly spaced monomers.

@@

  %BJC: old text:
%  When two vertically stacked monomer head groups hydrogen bond with a shared
%  water molecule, the monomers are drawn closer together (as illustrated in
%  Figure~\ref{fig:hbond_visualization}), which creates an asymmetry that allows
%  R-double to appear. If a monomer head group shares a hydrogen-bonded water
%  molecule with a head group above itself, it will be less likely to share a
%  water molecule with a head group below it due to geometric constraints. The
%  monomer head group below can just as easily share a water molecule with a head
%  group below itself. In this scenario, the COMs of each pair are 2 times the
%  $\pi$-stacking distance apart which would lead to R-double (much like the
%  configuration in Figure~\ref{fig:staggered_rzplot_norestraints}). There are a
%  modest number of occurrences of this scenario, which we quantify in further
%  detail in the SI, Section \ref{S-section:rdouble}.

  The addition of a small amount of water to the pores stabilizes non-uniform
  stacking of head groups much like that shown in
  Figure~\ref{fig:staggered_rzplot_norestraints}. When two vertically stacked
  monomer head groups hydrogen bond with a shared water molecule, the monomers
  are drawn closer together (as illustrated in
  Figure~\ref{fig:hbond_visualization}), which creates the asymmetry that allows
  R-double to appear. If a monomer head group shares a hydrogen-bonded water
  molecule with a head group above itself, it will be less likely to share a
  water molecule with a head group below it due to geometric constraints. The
  monomer head group below can just as easily share a water molecule with a head
  group below itself. There are a modest number of occurrences of this scenario,
  which we quantify in further detail in the SI, Section \ref{S-section:rdouble}.
